deprecated stuff that was tested.

## TESTING cluster log2FoldChanges from contrasts with first timepoint

This is a mess and needs attention. The fold changes have the wrong signs.

```{r}
# create data frame of contrasts between stage one and the others.
contr1 <- t(contr) %>% subset(.[,1]=="one") %>% t %>% as.data.frame(stringsAsFactors=FALSE)
# fix formatting/lines, rename call etc.
l2fChanges <- lapply(names(contr1), function(x) {results(dds, alpha=0.05, contrast=c("stage", contr1[,x]), tidy=TRUE) %>% select(row, log2FoldChange)})
# rename list elements to reflect contrast performed.
names (l2fChanges) <- lapply(names(contr1), function(x) {paste(contr1[,x], collapse="_vs_")}) %>% unlist
# create vector of contrast names for renaming l2fc columns.
contr2 <- lapply(names(contr1), function(x) {paste("l2fc", paste(contr1[,x], collapse="_vs_"), sep="_")}) %>% unlist
# rename l2fc columns in each list element.
for (i in seq_along(l2fChanges)) {
    names(l2fChanges[[i]]) <- c("row", contr2[i])
}
# merge all of the data frames in the list by gene name ("row").
resChanges <- Reduce(function(x, y) {merge(x, y, all=TRUE, by="row")}, l2fChanges)
# subset significant changes. Note for now this depends on the code chunks below but this could be refactored to include the lfc filter within the results calls on L84 but remember to modify the Reduce call with all=FALSE
resChangeLfc <- subset(resChanges, row %in% resLfc$row)
# write out for clustering.
write.table(select(resChangeLfc, -row), file="genes2", quote=FALSE, row.names=FALSE, sep=",")
```
Fit Dirichlet process mixture of Gaussians.

```{python}
    import numpy as np
    from sklearn import mixture

    # Load data without the headers.
    foo = np.loadtxt(open("genes2", "r"), delimiter=",", skiprows=1)

    n_components=40 # an upper bound for no. of components.
    lowest_bic = np.infty
    bic = []
    cv_types = ['spherical', 'tied', 'diag', 'full']
    for cv_type in cv_types:
        # Fit a Dirichlet process mixture of Gaussians.
        dpgmm = mixture.DPGMM(n_components=n_components, covariance_type=cv_type)
        dpgmm.fit(foo)
        bic.append(dpgmm.bic(foo))
        if bic[-1] < lowest_bic:
            lowest_bic = bic[-1]
            best_dpgmm = dpgmm

    # Predict the labels for the model with lowest BIC.
    labels = best_dpgmm.predict(foo)
    # Write the labels out for use in R.
    np.savetxt("labels2", labels, fmt="%s")
```

Read in labels, tidy, and plot by cluster.

```{r}
resChangeLfc$cluster <- scan("labels2", what="character")
resChangeLfcL <- gather(resChangeLfc, stage, expression, -row, -cluster) %>% 
	group_by(cluster, stage) %>%
	summarize(exp=mean(expression), sd=sd(expression))
# quick fix for wrong sign.
resChangeLfcL$exp <- resChangeLfcL$exp*-1
ggplot(resChangeLfcL, aes(x=stage, y=exp, group=cluster)) + geom_line() + facet_wrap(~cluster, ncol=5, scales="free_y") + ylab("Expression") + xlab("Stage") + ggtitle("Gene expression across larval-pupal moult") + geom_ribbon(aes(ymin=exp-sd, ymax=exp+sd), alpha=0.2)
```

## TESTING scale rlogged values

```{r}
scal_rld <- t(rldMat) %>% scale %>% t %>% as.data.frame %>% subset(rownames(.) %in% resLfc$row)
scal_rld$gene <- rownames(scal_rld)
colnames(scal_rld) <- c(paste(colData[,1] , colData$stage, sep="_"), "gene")
# Reshape into long format.
rsvt2 <- gather(scal_rld, stage, expression, -gene)
rsvt2$stage <- as.character(rsvt2$stage)
rsvt2 <- separate(rsvt2, stage, c("sample", "stage"))
by_genExp2 <- group_by(rsvt2, gene, stage)
# don't use devExp itself so will replace with a large chain
devExpCl2 <- summarize(by_genExp2, exp=mean(expression), sd=sd(expression))
devExpCw2 <- select(devExpCl2, gene, stage, exp) %>% spread(stage, exp) %>%
	select(gene, one, two, three, five, six, seven)
# nrow(res_rld)*6 == nrow(devExp) == nrow(devExpCw)*6
rownames(devExpCw2) <- devExpCw2$gene
```

```{r}
write.table(select(devExpCw2, -gene), file="genes3", quote=FALSE, row.names=FALSE, sep=",")
```

```{python}
import numpy as np
from sklearn import mixture

# Load data without the headers.
foo = np.loadtxt(open("genes3", "r"), delimiter=",", skiprows=1)

n_components=40 # an upper bound for no. of components.
lowest_bic = np.infty
bic = []
cv_types = ['spherical', 'tied', 'diag', 'full']
for cv_type in cv_types:
    # Fit a Dirichlet process mixture of Gaussians.
    dpgmm = mixture.DPGMM(n_components=n_components, covariance_type=cv_type)
    dpgmm.fit(foo)
    bic.append(dpgmm.bic(foo))
    if bic[-1] < lowest_bic:
        lowest_bic = bic[-1]
        best_dpgmm = dpgmm

# Predict the labels for the model with lowest BIC.
labels = best_dpgmm.predict(foo)
# Write the labels out for use in R.
np.savetxt("labels3", labels, fmt="%s")
```

```{r}
devExpCw2$cluster <- scan("labels3", what="character")
devExpCl3 <- gather(devExpCw2, stage, expression, -gene, -cluster) %>% 
	group_by(cluster, stage) %>%
	summarize(exp=mean(expression), sd=sd(expression))
ggplot(devExpCl3, aes(x=stage, y=exp, group=cluster)) + geom_line() + facet_wrap(~cluster, ncol=5, scales="free_y") + ylab("Expression") + xlab("Stage") + ggtitle("Gene expression across larval-pupal moult") + geom_ribbon(aes(ymin=exp-sd, ymax=exp+sd), alpha=0.2)
```

hmm, there are only two even clusters... sweeping by rowMeans lead to 16 clusters but with the majority of genes assigned to a single cluster.

Determine significantly loading genes from PCA.
```{r}
library(mseapca)
D <- data.frame(rownames(res_rld), select(res_rld, -gene))
A <- pca_scaled(D)
P <- A[["p.value"]]
colnames(P) <- c("gene", paste("PC", c(1:43), sep="_"))
Pl <- gather(P, loading, pc, -gene)
L <- A[["factor.loading"]]
Q <- A[["q.value"]]
colnames(Q) <- c("gene", paste("q_PC", c(1:43), sep="_"))

```

